---
title: "Final Project"
author: "Surendran"
date: '2022-06-28'
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(mice)
library(lattice)
library(dplyr)
library(magrittr)
library(tidyverse)
library(lme4)
library(broom)
library(VIM)
library(targets)
library(Metrics)
library(jtools)
library(zoo)
library(imputeTS)
library(xts)
setwd("C:/Users/suren/OneDrive - Robert Gordon University/Documents/R/RGU/C- Data visualisation & analysis/session  12 - TESTING")
```

```{r,warning=T, include=False}

if (!require('ggplot2')) install.packages(' ggplot2', dependencies = T);
if (!require('mice')) install.packages(' mice', dependencies = T);
if (!require('lattice')) install.packages(' lattice', dependencies = T);
if (!require('dplyr')) install.packages(' dplyr', dependencies = T);
if (!require('magrittr')) install.packages(' magrittr', dependencies = T);
if (!require('tidyverse')) install.packages(' tidyverse', dependencies = T);
if (!require('lme4')) install.packages(' lme4', dependencies = T);
if (!require('broom')) install.packages(' broom', dependencies = T);
if (!require('VIM')) install.packages(' VIM', dependencies = T);
if (!require('targets')) install.packages(' targets', dependencies = T);
if (!require('Metrics')) install.packages(' Metrics', dependencies = T);
if (!require('jtools')) install.packages(' jtools', dependencies = T);
if (!require('zoo')) install.packages(' zoo', dependencies = T);
if (!require('imputeTS')) install.packages(' imputeTS', dependencies = T);
if (!require('xts')) install.packages(' xts', dependencies = T);
```

                                                ORIGNAL DATASET

```{r}
dataset <- read.csv("dataset.csv",  header= T , stringsAsFactors = T )
summary(dataset)
head(dataset, 20)
```

# Plotting either missing dataset OR imputed-dataset by compiling each function

```{r}
library(VIM)
aggr_plot <- aggr(dataset1, col=c('green','red'),
                  numbers=TRUE,
                  sortVars=TRUE,
                  labels=names(dataset1),
                  cex.axis=.8,
                  gap=2,
                  xlab=c("Histogram of Missing or imputed data", "Pattern"))
                  
```

                                                MODIFIED DATASET
                                                

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

*************************IMPUTATION METHODS*************************************


# MEAN IMPUTATION

```{r}
for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}
```

# MEDIAN IMPUTATION

```{r}
for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}
```

# COMBINING ( LOCF + NOCB = IMPUTATION)

```{r}
dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))
```

# LINEAR INTERPOLATION IMPUTATION

```{r}
library(imputeTS)
dataset1 <- zoo(dataset1)

dataset1 <- t(dataset1)

dataset1 <- na.interpolation(dataset1, option = "linear")

dataset1 <- t(dataset1)

print(dataset1)
```

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

# Plotting either missing dataset or imputed-dataset before or after imputation

```{r}
library(VIM)
aggr_plot <- aggr(dataset1, col=c('green','red'),
                  numbers=TRUE,
                  sortVars=TRUE,
                  labels=names(dataset1),
                  cex.axis=.8,
                  gap=2,
                  xlab=c("Histogram of Missing or imputed data", "Pattern"))
                  
```

********************** EVALUATION OF TECHNIQUES **************

# CHOOSING 20 RANDOM EXISTING VALUES FOR EVALUATION PURPOSE

[1,1] [2,1] [3,2] [4,3] [5,5] [6,5] [7,1] [8,2] [9,3] [10,2]

[11,5] [12,1] [13,1] [14,1] [15,4] [16,5] [17,1] [18,2] [19,4] [20,1]

                                                MODIFIED DATASET
                                                

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

```{r}
differences <- c()
print(differences)
```

# EVALUATION for [1,1]

```{r}
original_value <- dataset1[1,1]

print(original_value)

dataset1[1,1] <- NA
```

```{r}
for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}
```

```{r}
imputed_value <- dataset1[1,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```





# EVALUATION for [1,1]

```{r}
original_value <- dataset1[1,1]

print(original_value)

dataset1[1,1] <- NA



for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[1,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [2,1]

```{r}
original_value <- dataset1[2,1]

print(original_value)

dataset1[2,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[2,1]
print(imputed_value)

difference = abs(original_value - imputed_value)


print(difference)


differences <- append(differences, difference)
print(differences)
```




# EVALUATION for [3,2]

```{r}
original_value <- dataset1[3,2]

print(original_value)

dataset1[3,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[3,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [4,3]

```{r}
original_value <- dataset1[4,3]

print(original_value)

dataset1[4,3] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[4,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [5,5]

```{r}
original_value <- dataset1[5,5]

print(original_value)

dataset1[5,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[5,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [6,5]

```{r}
original_value <- dataset1[6,5]

print(original_value)

dataset1[6,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[6,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [7,1]

```{r}
original_value <- dataset1[7,1]

print(original_value)

dataset1[7,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[7,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [8,2]

```{r}
original_value <- dataset1[8,2]

print(original_value)

dataset1[8,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[8,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [9,3]

```{r}
original_value <- dataset1[9,3]

print(original_value)

dataset1[9,3] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[9,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [10,2]

```{r}
original_value <- dataset1[10,2]

print(original_value)

dataset1[10,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[10,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [11,5]

```{r}
original_value <- dataset1[11,5]

print(original_value)

dataset1[11,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[11,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [12,1]

```{r}
original_value <- dataset1[12,1]

print(original_value)

dataset1[12,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[12,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [13,1]

```{r}
original_value <- dataset1[13,1]

print(original_value)

dataset1[13,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[13,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [14,1]

```{r}
original_value <- dataset1[14,1]

print(original_value)

dataset1[14,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[14,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [15,4]

```{r}
original_value <- dataset1[15,4]

print(original_value)

dataset1[15,4] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[15,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [16,5]

```{r}
original_value <- dataset1[16,5]

print(original_value)

dataset1[16,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[16,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [17,1]

```{r}
original_value <- dataset1[17,1]

print(original_value)

dataset1[17,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[17,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [18,2]

```{r}
original_value <- dataset1[18,2]

print(original_value)

dataset1[18,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[18,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [19,4]

```{r}
original_value <- dataset1[19,4]

print(original_value)

dataset1[19,4] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[19,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [20,1]

```{r}
original_value <- dataset1[20,1]

print(original_value)

dataset1[20,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[20,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```


# FINDING OVERALL MEAN FOR ALL OF THE ABOVE DIFFERENCES (MEAN IMPUTATION)

```{r}
mean(differences)
```















# MEDIAN imputation totally

```{r}
for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}
```

                                               DATASET

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

```{r}
differences <- c()
print(differences)
```

# EVALUATION for [1,1]

```{r}
original_value <- dataset1[1,1]

print(original_value)

dataset1[1,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[1,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [2,1]

```{r}
original_value <- dataset1[2,1]

print(original_value)

dataset1[2,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[2,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [3,2]

```{r}
original_value <- dataset1[3,2]

print(original_value)

dataset1[3,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[3,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [4,3]

```{r}
original_value <- dataset1[4,3]

print(original_value)

dataset1[4,3] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[4,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [5,5]

```{r}
original_value <- dataset1[5,5]

print(original_value)

dataset1[5,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}

imputed_value <- dataset1[5,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [6,5]

```{r}
original_value <- dataset1[6,5]

print(original_value)

dataset1[6,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[6,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [7,1]

```{r}
original_value <- dataset1[7,1]

print(original_value)

dataset1[7,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[7,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [8,2]

```{r}
original_value <- dataset1[8,2]

print(original_value)

dataset1[8,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[8,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [9,3]

```{r}
original_value <- dataset1[9,3]

print(original_value)

dataset1[9,3] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}

imputed_value <- dataset1[9,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [10,2]

```{r}
original_value <- dataset1[10,2]

print(original_value)

dataset1[10,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[10,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [11,5]

```{r}
original_value <- dataset1[11,5]

print(original_value)

dataset1[11,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[11,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [12,1]

```{r}
original_value <- dataset1[12,1]

print(original_value)

dataset1[12,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[12,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [13,1]

```{r}
original_value <- dataset1[13,1]

print(original_value)

dataset1[13,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[13,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [14,1]

```{r}
original_value <- dataset1[14,1]

print(original_value)

dataset1[14,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[14,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [15,4]

```{r}
original_value <- dataset1[15,4]

print(original_value)

dataset1[15,4] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[15,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [16,5]

```{r}
original_value <- dataset1[16,5]

print(original_value)

dataset1[16,5] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[16,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [17,1]

```{r}
original_value <- dataset1[17,1]

print(original_value)

dataset1[17,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[17,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [18,2]

```{r}
original_value <- dataset1[18,2]

print(original_value)

dataset1[18,2] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[18,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [19,4]

```{r}
original_value <- dataset1[19,4]

print(original_value)

dataset1[19,4] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[19,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [20,1]

```{r}
original_value <- dataset1[20,1]

print(original_value)

dataset1[20,1] <- NA


for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- median(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[20,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# FINDING OVERALL MEAN FOR ALL OF THE ABOVE DIFFERENCES (MEDIAN IMPUTATION)

```{r}
mean(differences)
```











# LAST OBSERVATION CARRIED FORWARD ( LOCF - from left to right)

```{r}
dataset1 <- data.frame(t(apply(dataset1, 1, na.locf0)))
```

# NEXT OBSERVATION CARRIED BACKWARD ( NOCB - from right to left)

```{r}
dataset1 <- data.frame(t(apply(dataset1, 1, na.locf0,fromLast = TRUE)))
```

# COMBINING ( LOCF + NOCB )

```{r}
dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))
```

                                               DATASET

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

```{r}
differences <- c()
print(differences)
```

# EVALUATION for [1,1]

```{r}
original_value <- dataset1[1,1]

print(original_value)

dataset1[1,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[1,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [2,1]

```{r}
original_value <- dataset1[2,1]

print(original_value)

dataset1[2,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[2,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [3,2]

```{r}
original_value <- dataset1[3,2]

print(original_value)

dataset1[3,2] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[3,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [4,3]

```{r}
original_value <- dataset1[4,3]

print(original_value)

dataset1[4,3] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[4,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [5,5]

```{r}
original_value <- dataset1[5,5]

print(original_value)

dataset1[5,5] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[5,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [6,5]

```{r}
original_value <- dataset1[6,5]

print(original_value)

dataset1[6,5] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[6,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [7,1]

```{r}
original_value <- dataset1[7,1]

print(original_value)

dataset1[7,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[7,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [8,2]

```{r}
original_value <- dataset1[8,2]

print(original_value)

dataset1[8,2] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[8,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [9,3]

```{r}
original_value <- dataset1[9,3]

print(original_value)

dataset1[9,3] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))

imputed_value <- dataset1[9,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [10,2]

```{r}
original_value <- dataset1[10,2]

print(original_value)

dataset1[10,2] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))


imputed_value <- dataset1[10,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [11,5]

```{r}
original_value <- dataset1[11,5]

print(original_value)

dataset1[11,5] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[11,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [12,1]

```{r}
original_value <- dataset1[12,1]

print(original_value)

dataset1[12,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[12,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [13,1]

```{r}
original_value <- dataset1[13,1]

print(original_value)

dataset1[13,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[13,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [14,1]

```{r}
original_value <- dataset1[14,1]

print(original_value)

dataset1[14,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[14,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [15,4]

```{r}
original_value <- dataset1[15,4]

print(original_value)

dataset1[15,4] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[15,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [16,5]

```{r}
original_value <- dataset1[16,5]

print(original_value)

dataset1[16,5] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[16,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [17,1]

```{r}
original_value <- dataset1[17,1]

print(original_value)

dataset1[17,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[17,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [18,2]

```{r}
original_value <- dataset1[18,2]

print(original_value)

dataset1[18,2] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[18,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [19,4]

```{r}
original_value <- dataset1[19,4]

print(original_value)

dataset1[19,4] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[19,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [20,1]

```{r}
original_value <- dataset1[20,1]

print(original_value)

dataset1[20,1] <- NA


dataset1 <- data.frame(t(apply(dataset1, 1, na.locf)))



imputed_value <- dataset1[20,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# FINDING OVERALL MEAN FOR ALL OF THE ABOVE DIFFERENCES( LOCF + NOCB IMPUTATION)

```{r}
mean(differences)
```








# LINEAR INTERPOLATION IMPUTATION

```{r}
library(imputeTS)
dataset1 <- zoo(dataset1)

dataset1 <- t(dataset1)

dataset1 <- na.interpolation(dataset1, option = "linear")

dataset1 <- t(dataset1)

print(dataset1)
```

                                               DATASET

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

```{r}
differences <- c()
print(differences)
```

# EVALUATION for [1,1]

```{r}
original_value <- dataset1[1,1]

print(original_value)

dataset1[1,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[1,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [2,1]

```{r}
original_value <- dataset1[2,1]
print(original_value)

dataset1[2,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[2,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [3,2]

```{r}
original_value <- dataset1[3,2]

print(original_value)

dataset1[3,2] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[3,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [4,3]

```{r}
original_value <- dataset1[4,3]

print(original_value)

dataset1[4,3] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[4,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [5,5]

```{r}
original_value <- dataset1[5,5]

print(original_value)

dataset1[5,5] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[5,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [6,5]

```{r}
original_value <- dataset1[6,5]

print(original_value)

dataset1[6,5] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[6,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [7,1]

```{r}
original_value <- dataset1[7,1]

print(original_value)

dataset1[7,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[7,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [8,2]

```{r}
original_value <- dataset1[8,2]

print(original_value)

dataset1[8,2] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[8,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [9,3]

```{r}
original_value <- dataset1[9,3]

print(original_value)

dataset1[9,3] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[9,3]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [10,2]

```{r}
original_value <- dataset1[10,2]

print(original_value)

dataset1[10,2] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[10,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [11,5]

```{r}
original_value <- dataset1[11,5]

print(original_value)

dataset1[11,5] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[11,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [12,1]

```{r}
original_value <- dataset1[12,1]

print(original_value)

dataset1[12,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[12,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [13,1]

```{r}
original_value <- dataset1[13,1]

print(original_value)

dataset1[13,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[13,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [14,1]

```{r}
original_value <- dataset1[14,1]

print(original_value)

dataset1[14,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[14,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [15,4]

```{r}
original_value <- dataset1[15,4]

print(original_value)

dataset1[15,4] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[15,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [16,5]

```{r}
original_value <- dataset1[16,5]

print(original_value)

dataset1[16,5] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[16,5]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [17,1]

```{r}
original_value <- dataset1[17,1]

print(original_value)

dataset1[17,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[17,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [18,2]

```{r}
original_value <- dataset1[18,2]

print(original_value)

dataset1[18,2] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[18,2]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```

# EVALUATION for [19,4]

```{r}
original_value <- dataset1[19,4]

print(original_value)

dataset1[19,4] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[19,4]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

```

# EVALUATION for [20,1]

```{r}
original_value <- dataset1[20,1]

print(original_value)

dataset1[20,1] <- NA


library(imputeTS)
dataset1 <- zoo(dataset1)
dataset1 <- t(dataset1)
dataset1 <- na.interpolation(dataset1, option = "linear")
dataset1 <- t(dataset1)
print(dataset1)


imputed_value <- dataset1[20,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```



# FINDING OVERALL MEAN FOR ALL OF THE ABOVE DIFFERENCES (LINEAR INTERPOLATION)

```{r}
mean(differences)
```

















# LINEAR INTERPOLATION MANUAL DISPLAY INSTEAD OF AUTOMATIC-FLLING

```{r}
imputeTS::na_interpolation(dataset1)
```

# 06/07/2022

# TO REMOVE COMPLETE ROWS IN SPECIFIC ( below code deletes row# 2 & 4 from dataset1 )

```{r}
dataset1 <- slice(dataset1, -c(2, 4))
```

# TO REMOVE SPECIFIC COLUMNS ( below code deletes column# 1 & 3 from dataset1 )

```{r}
dataset1$Year2000 <- NULL
dataset1$Year2001 <- NULL
```

```{r}
dataset1[ 3,2] <- NA
```

ORIGNAL VALUE FROM 11TH ROW & 20TH COLUMN = 16.0150779

Mean = 17.5886053 , Median = 16.8453771 , LOCF = 15.4067261 , Linear
interpolation = 13.9530922

diff = 1.57(SECOND) , = 0.83 , = 0.61 , = 2.06 (LAST)

                                               DATASET

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

                                                ORIGNAL DATASET

```{r}
dataset <- read.csv("dataset.csv",  header= T , stringsAsFactors = T )
summary(dataset)
head(dataset, 20)
```

# OBSERVATION CARRIED BACKWARD (from down to up)

```{r}
dataset = lapply(dataset, zoo::na.locf, fromLast = TRUE)
```

------------------------------------------------------------------------

```{r}
differences <- c()
print(differences)
```

```{r}

original_value <- dataset1[1,1]
print(original_value)

original_value <- dataset1[2,1]
print(original_value)

dataset1[1,1] <- NA
dataset1[2,1] <- NA



for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}

imputed_value <- dataset1[1,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)

imputed_value <- dataset1[2,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)

differences <- append(differences, difference)
print(differences)
```

                                              ***NEW***

# MEAN imputation totally

```{r}
for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}
```

```{r}
differences <- c()
print(differences)
```

```{r}
dataset1 <- read.csv("dataset1.csv",  header= T , stringsAsFactors = T )
summary(dataset1)
head(dataset1, 20)
```

print(iris)

my_list \<- list(c(1,2),c(2,3),c(3,4)) for (my_item in my_list) { row
\<- my_item[1] col \<- my_item[2] print(iris[row, col]) }

# EVALUATION for [1,1]

```{r}
original_value <- dataset1[1,1]

print(original_value)

dataset1[1,1] <- NA



for(i in 1:nrow(dataset1)) 
  {
  dataset1[i, ][is.na(dataset1[i, ])] <- mean(as.numeric(dataset1[i, ]), na.rm = TRUE)
}


imputed_value <- dataset1[1,1]
print(imputed_value)

difference = abs(original_value - imputed_value)
print(difference)


differences <- append(differences, difference)
print(differences)
```
